---
title: "Exploring and visualizing results"
output: 
  html_document:
    toc: true
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

IN PROGRESS

# rvatResult

The `rvatResult()` class is designed to work with results from RVAT
analyses, and extends the `DataFrame` class (see:
[S4Vectors](https://bioconductor.org/packages/devel/bioc/manuals/S4Vectors/man/S4Vectors.pdf)).
It supports all kinds of common operations such as subsetting and
merging, and several methods are defined for reading&writing and
(interactive) visualization.

## Reading & Writing

The `rvbResult()` function loads a results file written to disk (note
that for single variant results there is an equivalent
`singlevarResult()`) function:`

```{r, message = FALSE, warning = FALSE}
library(rvat)
library(rvatData)
outdir <- tempdir()

# Load example results file
data(rvbresults)
head(rvbresults)
```

The `writeResult()` method writes the data to disk:

```{r}
writeResult(rvbresults, file = paste0(outdir,"/burden_results.txt.gz"))
```

The `rvbResult()` function can be used to load results:

```{r}
rvbresults <- rvbResult(paste0(outdir,"/burden_results.txt.gz"))
```

## Inspecting the results

The `summary()` method gives a summary of the contents of the results
file:

```{r}
summary(rvbresults)
```

Several methods such as `head()`, `dim()`, `ncol()` and `nrow()` can be
used to inspect the results object:

```{r}
rvbresults[1:5,]
```

```{r}
dim(rvbresults)
```

```{r}
ncol(rvbresults)
```

```{r}
head(rvbresults)
```

subsetting is identical to data.frame:

```{r}
rvbresults[rvbresults$varSetName == "ModerateImpact" & rvbresults$test == "firth", ]
```

## Combining, merging & splitting

Results files can be combined using `rbind()`:

```{r}
# for this example we'll split the results-file in two and then glue it back together:
results1 <- rvbresults[1:ceiling(nrow(rvbresults)/2), ]
results2 <- rvbresults[(ceiling(nrow(rvbresults)/2)+1):nrow(rvbresults), ]
test <- rbind(results1, results2)

# check if it's identical:
identical(rvbresults, test)
```

add additional info to the results using `merge()`, where the `by` parameter
specifies which columns should be used to join:

```{r, message = FALSE, warning = FALSE}
annotation <- as.data.frame(rvbresults)[,colnames(rvbresults) %in% c("unit", "CHROM","POS")]
annotation <- annotation[!duplicated(annotation$unit),]
rvbresults <- merge(rvbresults[,!colnames(rvbresults) %in% c("CHROM","POS")], 
                 annotation,
                 by = "unit")
```

the `split()` method can be used to split a results object into defined
groups. The example below showcases how to split the results into three
result objects (returned in a list) based on the `varSetName` column:

```{r}
results_list <- S4Vectors::split(rvbresults, as.character(rvbresults$varSetName))
head(results_list[[1]])
head(results_list[[2]])
```

# Visualize

## QQ-plot

```{r}
qqplot(rvbresults[rvbresults$test == "firth" &
       rvbresults$varSetName == "ModerateImpact",])
```

## Manhattan

Manhattan plots can be generated using the `manhattan()` method. The
`label` parameter can be specified to indicate a column with labels that
should be displayed for significant results:

```{r}
manhattan(rvbresults[rvbresults$test == "skat_burden_robust" &
                  rvbresults$varSetName == "ModerateImpact",], 
          label = "unit", 
          contigs = "GRCh38")
```

## rvatViewer

The `rvatViewer()` method allows for interactive exploration of the
results, it can be launched as following:

```{r, eval = FALSE}
rvatViewer(rvbresults)
```

A video tutorial on using `rvatViewer()` will be included in the near
future!

# References
